Effects of Antenna Beam Chromaticity on Redshifted 21~cm Power Spectrum and Implications for Hydrogen Epoch of Reionization Array

### 2016-05-06 ForegroundSimulations/reply_to_referee_report.txt

While 21cm cosmology promises a tremendous amount of new information in both astrophysics and cosmology, the weakness of the signal combined with the dominance of foregrounds poses daunting requirements on instrument characterization. The work presented in this manuscript directly addresses some of the most urgent yet poorly understood aspects of 21 cm instruments, so it is highly valuable to the field and timely for the deployment and future expansion of the HERA project. Therefore, I recommend this work for publication, with minor modifications to address my comments below:

Major comments:

1) It is not clear if the design of HERA dish is good enough or not. In sections 2 and 6, the authors convey the message that HERA dish design is good enough (even "optimal"), whereas in section 5 it is clear that more dish measurements and possible design tweaks are needed. It is important to present a more coherent picture on this front.

Response: The intention of this paper and companion papers is to illustrate that HERA is basing its design after comprehensive analysis of its effects on detectability of the EoR signal. Though many parameters of the HERA design are close to being final, efforts are ongoing to fine-tune the instrument based on measurements. I have rephrased statements to emphasize this picture. It must also be noted that the dish design is close to final while efforts are ongoing to obtain an optimal feed design.

@@

Response: I have noted in section 5.1 that foreground contamination depends on baseline length, orientation, pointing direction and spatial structure of emitting source. Hence, I chose a range of baseline lengths and orientations. I used all unique baseline lengths to demonstrate that all baselines individually will be sensitive to the EoR signal.

Minor comments:

Abstract:
1) "with high significance": this sounds vague. Is there a sigma value or a signal-to-noise ratio (SNR) to quote?

Response: Unless thermal noise is included it is difficult to quote a SNR. Here I have assumed data is not thermal noise limited but only limited by foregrounds and systematics studied in this paper. In such a scenario the significance of detection again varies anywhere from 10 to more than 1000. The exact detection significance will depend on the analysis methods and weighting schemes used.

@@

2) "All baselines...are capable": does this mean each baseline individually can make the detection, or all of them combined? Since this work focuses on beam chromaticity and just assumes sufficient SNR, it is not clear if such statement regarding detection is truly justifiable in the scope of this work.

Response: It is meant to imply all baselines individually can make the detection as long as they are not thermal-noise limited but only limited by foreground and instrumental systematics studied here. Hence the sentence is now prefixed/qualified with "Under these conditions" which refers to data that is limited by foregrounds and such systematics discussed in this paper.

@@

1) "currently under construction on the South African SKA site": nit-picking here (everywhere?), but it might sound confusing why HERA is built on the SKA site. Maybe just omit the SKA site information since it is not relevant?

Response: Replaced SKA site with "South African Karoo Radio Astronomy Reserve".

2) "strike an optimal balance between sensitivity and systematics": this is directly related to my major comments #1. First of all, even if that were the case, it is not quite appropriate to state it as a matter of fact in the introduction, since it is what this work is setting out to prove. Secondly, it seems clear in this work that the dish design has not truly achieved the optimum. Sections 5.2.1 and 5.2.2 explicitly state that 'measurements are under way" and "the HERA collaboration constantly pursues improvement of its dish design".

Response: Removed the word "optimal". The balance between sensitivity and systematics was initially studied in DeBoer and Parsons (2015) - a HERA memo - to which a citation has now been added. This study was specific to choice of dish size and preliminary estimates of specifications for dish based on sensitivity and systematics. Our current work along with companion papers are a much more comprehensive effort to further reinforce findings of that preliminary study. It is in this context that sentence was used. Unlike the dish design, it is true that the feed design is not finalized yet and is still being fine-tuned for optimum performance. Hence, removal of the word "optimum" is consistent with other statements about measurements and fine-tuning of the HERA element (dish and feed together) design.

@@

2) The authors go into great detail as to what the materials are used in simulation, which leaves the reader wondering: are they the same as what are actually used in the dish? For example, "the backplane is modeled as an aluminum surface", but is it actually an aluminum surface? A nice picture of an actual HERA dish may help here.

Response: TBD

3) Can the authors briefly comment on the relative importance of point sources versus the diffuse emission in the simulation? Are the point sources relevant for the 14 m baselines? What about the 60 m ones?

Response: For the HERA-19 layout the diffuse emission has a dominant effect. This can be seen from how the foreground emission peak drops with increasing baseline length and how the dynamic range in figure 6 increases with baseline length. I have added a sentence in the penultimate paragraph of section 6.2 to highlight this.

### 2016-05-06 ForegroundSimulations/ms.tex
Address relative importance of emission type
From 31e2ef7 (fgsim):
====================

In ms.tex,

(*) Change "dish design" to "element design"
(*) Add some sentences to indicate relative significance of diffuse
and point source emission

In reply_to_referee_report.txt,

(*) Address referee comment on relative significance of diffuse and
point source emission

In this paper, we investigate the spectral properties of the proposed dish design from a foreground contamination standpoint and the constraints they place on attenuation required to suppress reflections in the instrument.

@@

Increase in beam chromaticity makes the lower limits on attenuating reflections more severe relative to that from an achromatic beam. For instance, for $k_\textrm{p}=0.1\,h$~Mpc$^{-1}$ there is no delay at which the required attenuation is unconstrained to the left of the elbow-shaped turnover, including at $\tau=0$. This means that when additional chromaticity from antenna-to-antenna reflections is taken into account, the requirement that all $k_\parallel$-modes with $k_\parallel > 0.1\,h$~Mpc$^{-1}$ be accessible for EoR signal detection will not be satisfied with the currently simulated chromatic beam for HERA. For $k_\textrm{p}=0.15\,h$~Mpc$^{-1}$, there is a narrow range of allowed attenuation which is unconstrained for $\tau \lesssim 50$~ns. However, for $k_\parallel > 0.2\,h$~Mpc$^{-1}$, the foreground power spectrum using simulated chromatic dish beam will have enough room to be effective for EoR signal detection despite additional chromaticity arising from antenna-to-antenna reflections.

We note here that designing a dish whose antenna beam closely resembles a nominal {\it Airy} pattern will have a significantly higher tolerance for allowing antenna-to-antenna reflections and yet remain very effective for EoR signal detection. The HERA collaboration constantly pursues improvement of its dish design to minimize limitations from such chromatic systematics.

@@

We define baseline-dependent $k_\parallel$-modes of interest for {\it foreground avoidance} as $|k_\parallel| > k_\parallel^\textrm{FA}(\boldsymbol{b})$, with:
\begin{align}
  k_\parallel^\textrm{FA}(\boldsymbol{b}) &= \frac{2\pi\,f_{21}H_0\,E(z)}{c(1+z)^2}\,\left(\frac{|\boldsymbol{b}|}{c} + \frac{\zeta}{B_\textrm{eff}}\right), \label{eqn:kprll-limit}
\end{align}
where, the first term inside the parenthesis denotes the horizon limit, the second relates to the resolution due to subband bandwidth, $B_\textrm{eff}$, and $\zeta$ denotes a buffer to ensure the $k_\parallel^\textrm{FA}(|\boldsymbol{b}|)$ threshold ``safely'' avoids the main lobe of foreground power. In this study, we use $\zeta=3.5$.

@@

For the baselines considered, $\rho(\boldsymbol{b},f_0,\alpha)$ increases with baseline length. We attribute this to the much steeper dependence of foreground emission (predominantly diffuse emission) on baseline length than that of EoR H{\sc i} power. This is also confirmed from Fig.~\ref{fig:subbands} where the peak and sidelobe levels of foreground power obtained with the simulated chromatic antenna beam on the 14.6~m baseline, progressively drops by two orders of magnitude on the 58.4~m baseline. Hence, from a foreground contamination standpoint, the shortest baselines are less sensitive than longer ones in HERA-19 layout.

However, in all four cases considered here, except when the Galactic {\bf plane} appears overhead {\bf roughly between 16 and 19 hours of RA}, the worst case signal-foreground ratio lies above unity in most of this parameter space. It demonstrates that HERA should detect the EoR signal with a very high signal-foreground ratio on all baselines over a majority of the pointings.

>>>>

In this paper, we investigate the spectral properties of the proposed HERA {\bf element (dish and feed)} design from a foreground contamination standpoint and the constraints they place on attenuation required to suppress reflections in the instrument.

@@

Increase in beam chromaticity makes the lower limits on attenuating reflections more severe relative to that from an achromatic beam. For instance, for $k_\textrm{p}=0.1\,h$~Mpc$^{-1}$ there is no delay at which the required attenuation is unconstrained to the left of the elbow-shaped turnover, including at $\tau=0$. This means that when additional chromaticity from antenna-to-antenna reflections is taken into account, the requirement that all $k_\parallel$-modes with $k_\parallel > 0.1\,h$~Mpc$^{-1}$ be accessible for EoR signal detection will not be satisfied with the currently simulated chromatic beam for HERA. For $k_\textrm{p}=0.15\,h$~Mpc$^{-1}$, there is a narrow range of allowed attenuation which is unconstrained for $\tau \lesssim 50$~ns. However, for $k_\parallel > 0.2\,h$~Mpc$^{-1}$, the foreground power spectrum using simulated chromatic beam will have enough room to be effective for EoR signal detection despite additional chromaticity arising from antenna-to-antenna reflections.

We note here that designing {\bf an antenna element whose beam} closely resembles a nominal {\it Airy} pattern will have a significantly higher tolerance for allowing antenna-to-antenna reflections and yet remain very effective for EoR signal detection. The HERA collaboration constantly pursues improvement of its {\bf antenna element} design to minimize limitations from such chromatic systematics.

@@

We define baseline-dependent $k_\parallel$-modes of interest for {\it foreground avoidance} as $|k_\parallel| > k_\parallel^\textrm{FA}(\boldsymbol{b})$, with:
\begin{align}
  k_\parallel^\textrm{FA}(\boldsymbol{b}) &= \frac{2\pi\,f_{21}H_0\,E(z)}{c(1+z)^2}\,\left(\frac{|\boldsymbol{b}|}{c} + \frac{\zeta}{B_\textrm{eff}}\right), \label{eqn:kprll-limit}
\end{align}
where, the first term inside the parenthesis denotes the horizon limit, the second relates to the resolution due to subband bandwidth, $B_\textrm{eff}$, and $\zeta$ denotes a buffer to ensure the $k_\parallel^\textrm{FA}(|\boldsymbol{b}|)$ threshold ``safely'' avoids the main lobe of foreground power \citep{thy13}. In this study, we use $\zeta=3.5$.

@@

For the baselines considered, $\rho(\boldsymbol{b},f_0,\alpha)$ increases with baseline length. We attribute this to the much steeper dependence of foreground emission on baseline length than that of EoR H{\sc i} power. {\bf This trend demonstrates that it is predominantly due to diffuse emission}. This is also confirmed from Fig.~\ref{fig:subbands} where the peak and sidelobe levels of foreground power obtained with the simulated chromatic antenna beam on the 14.6~m baseline, progressively drops by two orders of magnitude on the 58.4~m baseline. Hence, from a foreground contamination standpoint, the shortest baselines are less sensitive than longer ones in HERA-19 layout. % {\bf With the full proposed layout emission from point sources may become significant on longer baselines}.

In all four cases considered here, except when the Galactic {\bf plane} appears overhead {\bf roughly between 16 and 19 hours of RA}, the worst case signal-foreground ratio lies above unity in most of this parameter space. It demonstrates that HERA should detect the EoR signal with a very high signal-foreground ratio on all baselines over a majority of the pointings.
